Values disclosures and trust in science: A replication study

### 2022-10-12 paper/paper.Qmd
finished revisions per reviewers

While this kind of argument indicates that at least some social and political values may or should play at least some role in the evaluation of at least some scientific hypotheses, it doesn't provide us with much positive guidance:  which values, playing which roles, in the evaluation of which hypotheses?  [@HicksNewDirectionScience2014]  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research [@ElliottSciencePolicyTransparency2014; @McKaughanBacktrackingEthicsFraming2013].  But some philosophers have objected to this transparency proposal, arguing that it might actually undermine trust in science.  According to this objection, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021].

@@

Ethicists make a key distinction between trust and trustworthiness [@BaierTrustAntitrust1986].  Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.  Trustworthiness is an assessment, of whether or not that trust is appropriate.  In the experiment, participants might trust Dr. Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa; though we would typically expect these two to go together.  As its name indicates, METI assesses perceived trustworthiness, not trust (such as accepting Dr. Spence's conclusion about BPA).

@@

@ElliottValuesEnvironmentalResearch2017 claim evidence for a shared values effect by comparing values disclosure vs. no disclosure (that is, estimating a transparency effect) in quasi-independent subsamples across a 2 $\times$ 2 $\times$ 2 design (participant values $\times$ scientist conclusion $\times$ scientist values; @ElliottValuesEnvironmentalResearch2017 fig. 1).  This approach is difficult to interpret; their reasoning seems to be that this estimate is statistically significant in 4/8 cells, and in 3/4 of these cells (1 where respondent values economic growth) the participant and scientist share values.  But these analyses are underpowered (for example, sample sizes are well below 100 for some of the cells in which participants value economic growth) and the overall approach cannot distinguish a shared values effect from other potential effects (scientist values, participant values).  Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$ = 0.29, *t*(335) = 1.75, *p* = 0.08; while the estimate for scientist values was statistically significant, $\beta$ = 0.41, *t*(335) = 2.48, *p* = 0.014.  It seems likely that claims of a shared values effect could have been the result of a less appropriate analytic approach given the nature of the data.

The disagreement over a transparency penalty is more difficult to explain.  Using our approach and the data from @ElliottValuesEnvironmentalResearch2017, the estimate for a values disclosure was statistically significant (@tbl-bc model 1).  One possible explanation is that, over the course of the COVID-19 pandemic, members of the general public have become used to "scientists" (including both bench researchers and public health officials) making claims about the importance of protecting public health.  A values disclosure that might have been regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.

@@

>>>>

While this kind of argument indicates that at least some social and political values may or should play at least some role in the evaluation of at least some scientific hypotheses, it doesn't provide us with much positive guidance:  which values, playing which roles, in the evaluation of which hypotheses?  [@HicksNewDirectionScience2014]  One (partial) answer to this set of questions appeals to the ideal of transparency:  scientists should disclose to the public the social and political values that have influenced their research [@ElliottSciencePolicyTransparency2014; @McKaughanBacktrackingEthicsFraming2013].  But some philosophers have objected to this transparency proposal, arguing that it might undermine trust in science.

Specifically, ethicists make a distinction between trust and trustworthiness [@BaierTrustAntitrust1986].  Trust is a verb: placing one's trust in someone else, with respect to some activity or domain.  Trustworthiness is an assessment, of whether or not that trust is appropriate.  According to the objection to the ideal of transparency, transparency would not change how trustworthy science actually is — the science is conducted in the same way — but might reduce public perceptions of trustworthiness — as members of the public see how the scientific sausage is made.  For instance, members of the public generally accept the ideal of value-free science; values disclosures would violate this ideal, making scientists (incorrectly) appear biased and untrustworthy [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021].

@@

As its name indicates, METI assesses perceived trustworthiness rather than trust (such as accepting Dr. Spence's conclusion about BPA).  It is conceivable that participants might trust Dr. Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa.  But we would typically trust and trustworthiness to go together.

@@

@ElliottValuesEnvironmentalResearch2017 claim evidence for a shared values effect by comparing values disclosure vs. no disclosure (that is, estimating a transparency effect) in quasi-independent subsamples across a 2 $\times$ 2 $\times$ 2 design (participant values $\times$ scientist conclusion $\times$ scientist values; @ElliottValuesEnvironmentalResearch2017 fig. 1).  This approach is difficult to interpret; their reasoning seems to be that this estimate is statistically significant in 4/8 cells, and in 3/4 of these cells (1 where respondent values economic growth) the participant and scientist share values.  But these analyses are underpowered (for example, sample sizes are well below 100 for some of the cells in which participants value economic growth) and the overall approach cannot distinguish a shared values effect from other potential effects (scientist values, participant values).  Applying our regression analysis approach to the data published by @ElliottValuesEnvironmentalResearch2017, the estimate for shared values was not statistically significant, $\beta$ = 0.29, *t*(335) = 1.75, *p* = 0.08; while the estimate for scientist values was statistically significant, $\beta$ = 0.41, *t*(335) = 2.48, *p* = 0.014.  It seems likely that claims of a shared values effect could have been an artifact of a less appropriate analytic approach given the nature of the data.

The disagreement over a transparency penalty is more difficult to explain.  Using our approach and the data from @ElliottValuesEnvironmentalResearch2017, the estimate for a values disclosure was statistically significant (@tbl-bc model 1).  One possible explanation is that, over the course of the COVID-19 pandemic, members of the general public have become used to "scientists" (broadly including both bench researchers and public health officials) making claims about the importance of protecting public health.  A values disclosure that might have been regarded as violating the value-free ideal, pre-pandemic, might now be seen as routine.

@@

All together, regarding the philosophical debates over values in science, the evidence from this replication study suggests that the effects of transparency on perceived trustworthiness might be context-dependent.  Transparency about the influence of might be detrimental in some cases, beneficial in others, and neutral in still others.  Notably, both of the papers cited as objecting to transparency [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021] focus on climate change, while many of Elliott's examples come from toxicology or environmental public health.  The critics might be right that transparency has undermined trust in the climate context, while Elliott might be right that transparency can be neutral or even beneficial in some public health contexts.  Because philosophers of science tend to work with a small number of rich case studies, methods such as qualitative comparative analysis [QCA\; @RaginComparativeMethodMoving2013] would be worth exploring.

An obvious difference between climate and BPA is the public prominence and political polarization of the two issues.  For decades, the fossil fuels industry has used a range of strategies to successfully attack the perceived trustworthiness of climate scientists, especially in the minds of conservative publics [@OreskesMerchantsDoubtHow2011; @SupranAssessingExxonMobilClimate2017].  While the chemical industry has also had a substantial influence on the environmental safety regulatory system [@MichaelsDoubtTheirProduct2008; @VogelItSafeBPA2013], controversies over BPA have received much less public attention.  A LexiNexis search indicates that uses of "bisphenol" in US newspapers rose and then almost immediately fell around 2008, shortly after a US consensus conference concluded that "BPA at concentrations found in the human body is associated with organizational changes" in a number of major physiological systems [@vomSaalChapelHillBisphenol2007].

In more prominent and/or polarized cases than BPA, a number of other psychological mechanisms might be activated.  Motivated reasoning or identity-protective cognition [@KahanPolarizingImpactScience2012] might cause members of the general public to rate a scientist as more or less trustworthy insofar as the scientist agrees with the respondent's "side" of the debate, regardless of whether and which values are disclosed.  That is, the scientist conclusion might have a larger effect than scientist values.  Alternatively, respondents might use values disclosures to identify which scientists are on "their side," and so scientist values would have a larger effect than scientist conclusion.  Careful attention to experimental design will be needed to separate these potential multiway interactions between conclusions, disclosures, and public prominence.

### 2022-10-12 paper/paper.tex
finished revisions per reviewers

\begin{document}
\maketitle
\begin{abstract}
While philosophers of science generally agree that social, political,
and ethical values can play legitimate roles in science, there is active
debate over whether scientists should disclosure such values in their
public communications. This debate depends, in part, on empirical claims
about whether values disclosures might undermine public trust in
science. In a previous study, Elliott et al.~used an online experiment
to test this empirical claim. The current paper reports a replication
attempt of their experiment. Comparing results of the original study and
our replication, we do not find evidence for a transparency penalty or
``shared values'' effect, but do find evidence that the content of
scientific conclusions (whether or not a chemical is found to cause
harm) might effect perceived trustworthiness and that scientists who
value public health and disclose this value might be perceived as more
trustworthy.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

While this kind of argument indicates that at least some social and
political values may or should play at least some role in the evaluation
of at least some scientific hypotheses, it doesn't provide us with much
positive guidance: which values, playing which roles, in the evaluation
of which hypotheses? (Hicks 2014) One (partial) answer to this set of
questions appeals to the ideal of transparency: scientists should
disclose to the public the social and political values that have
influenced their research (Elliott and Resnik 2014; McKaughan and
Elliott 2013). But some philosophers have objected to this transparency
proposal, arguing that it might actually undermine trust in science.
According to this objection, members of the public generally accept the
ideal of value-free science; values disclosures would violate this
ideal, making scientists (incorrectly) appear biased and untrustworthy
(John 2017; Kovaka 2021).

@@

After viewing the slide, the participant is then asked to rate
Dr.~Spence's trustworthiness using a semantic differential scale.
Elliott et al. (2017) used an ad hoc scale; we used the Muenster
Epistemic Trustworthiness Inventory (METI; Hendriks, Kienhues, and
Bromme 2015), which substantially overlaps but is not identical to the
Elliott et al. (2017) scale. On the METI, participants rate the target
scientist (the fictional Dr.~Spence) on a 1-7 scale for 14 semantic
differential items, where each item is anchored at the ends by a pair of
words, such as competent-incompetent or responsible-irresponsible. The
METI scale captures perceived trustworthiness of a target along three
dimensions: competence, benevolence, and integrity. However, the three
dimensions were strongly correlated in our sample. To avoid issues of
multicollinearity in our analyses, we averaged participant scores into a
single composite measure. In both Elliott et al. (2017) and our
replication analysis, all items were averaged together to form a 1-7
composite measure of trustworthiness. To aid interpretation, in the
analysis the direction of the scale has been set so that increased
values correspond to increased perceived trustworthiness.

Ethicists make a key distinction between trust and trustworthiness
(Baier 1986). Trust is a verb: placing one's trust in someone else, with
respect to some activity or domain. Trustworthiness is an assessment, of
whether or not that trust is appropriate. In the experiment,
participants might trust Dr.~Spence's conclusion even if they judge
Spence to be untrustworthy, or vice versa; though we would typically
expect these two to go together. As its name indicates, METI assesses
perceived trustworthiness, not trust (such as accepting Dr.~Spence's
conclusion about BPA).

@@

\begin{description}
\tightlist
\item[H1. Modest correlation between values and ideology]
(a) Political liberals are more likely to prioritize public health over
economic growth, compared to political conservatives; but (b) a majority
of political conservatives prioritize public health.
\item[H2. Consumer risk sensitivity]
Scientists who find that a chemical harms human health are perceived as
more trustworthy than scientists who find that a chemical does not cause
harm.
\item[H3. Transparency penalty]
Scientists who disclose values are perceived as less trustworthy than
scientists who do not.
\item[H4. Shared values]
Given that the scientist discloses values, if the participant and the
scientist share the same values, the scientist is perceived as more
trustworthy than if the participant and scientist have discordant
values.
\item[H5. Variation in effects]
The magnitude of the effects 2-4 vary depending on whether the
participant prioritizes public health or economic growth.
\end{description}

@@

Hypothesis 1 was analyzed using Spearman rank order correlation and
visually. Hypotheses 2-5 were analyzed using linear regression models as
a common framework, with a direct acyclic graph (DAG) constructed
\emph{a priori} to identify appropriate adjustments (covariates) for H4
and 5. Because Elliott et al. (2017) made their data publicly available,
and our analytical approaches differ somewhat from theirs, we conducted
parallel analyses of their data, and report these parallel results for
H2 and 3. Exploratory data analysis was used throughout to support data
validation and aid interpretation.

@@

After excluding participants who declined consent after opening the
survey or did not complete the survey, we had 988 participants in the
full analysis sample (\(M_{age}\) = 44-years-old, \(SD_{age}\) =
16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124,
Asian or Pacific Islander = 63, Hispanic = 33, American Indian or
Alaskan Native = 5, Mixed or Other = 51). Participants were randomly
assigned to condition, with 163 assigned to the No Disclosure + Causes
Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm
condition, 165 assigned to the Economic Growth + Causes Harm condition,
165 assigned to the Economic Growth + Does Not Cause Harm condition, 168
assigned to the Public Health + Causes Harm condition, and 162 assigned
to the Public Health + Does Not Cause Harm condition
(Table~\ref{tbl-condition}). Of the full 988 participants, 844
participants (85\%) responded to the followup question about their own
values (participant prioritizes economic growth or public health).
Consequently, subsamples for hypotheses 4 and 5 were substantially
smaller than the full analysis sample.

@@

Elliott et al. (2017) claim evidence for a shared values effect by
comparing values disclosure vs.~no disclosure (that is, estimating a
transparency effect) in quasi-independent subsamples across a 2
\(\times\) 2 \(\times\) 2 design (participant values \(\times\)
scientist conclusion \(\times\) scientist values; Elliott et al. (2017)
fig.~1). This approach is difficult to interpret; their reasoning seems
to be that this estimate is statistically significant in 4/8 cells, and
in 3/4 of these cells (1 where respondent values economic growth) the
participant and scientist share values. But these analyses are
underpowered (for example, sample sizes are well below 100 for some of
the cells in which participants value economic growth) and the overall
approach cannot distinguish a shared values effect from other potential
effects (scientist values, participant values). Applying our regression
analysis approach to the data published by Elliott et al. (2017), the
estimate for shared values was not statistically significant, \(\beta\)
= 0.29, \emph{t}(335) = 1.75, \emph{p} = 0.08; while the estimate for
scientist values was statistically significant, \(\beta\) = 0.41,
\emph{t}(335) = 2.48, \emph{p} = 0.014. It seems likely that claims of a
shared values effect could have been the result of a less appropriate
analytic approach given the nature of the data.

The disagreement over a transparency penalty is more difficult to
explain. Using our approach and the data from Elliott et al. (2017), the
estimate for a values disclosure was statistically significant
(Table~\ref{tbl-bc} model 1). One possible explanation is that, over the
course of the COVID-19 pandemic, members of the general public have
become used to ``scientists'' (including both bench researchers and
public health officials) making claims about the importance of
protecting public health. A values disclosure that might have been
regarded as violating the value-free ideal, pre-pandemic, might now be
seen as routine.

@@

@@

@@

@@

@@

@@

@@

>>>>

\begin{document}
\maketitle
\begin{abstract}
While philosophers of science generally agree that social, political,
and ethical values can play legitimate roles in science, there is active
debate over whether scientists should disclosure such values in their
public communications. This debate depends, in part, on empirical claims
about whether values disclosures might undermine public trust in
science. In a previous study, Elliott et al.~used an online experiment
to test this empirical claim. The current paper reports a replication
attempt of their experiment. Comparing results of the original study and
our replication, we do not find evidence for a transparency penalty or
``shared values'' effect, but do find evidence that the content of
scientific conclusions (whether or not a chemical is found to cause
harm) might effect perceived trustworthiness and that scientists who
value public health and disclose this value might be perceived as more
trustworthy.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

While this kind of argument indicates that at least some social and
political values may or should play at least some role in the evaluation
of at least some scientific hypotheses, it doesn't provide us with much
positive guidance: which values, playing which roles, in the evaluation
of which hypotheses? (Hicks 2014) One (partial) answer to this set of
questions appeals to the ideal of transparency: scientists should
disclose to the public the social and political values that have
influenced their research (Elliott and Resnik 2014; McKaughan and
Elliott 2013). But some philosophers have objected to this transparency
proposal, arguing that it might undermine trust in science.

Specifically, ethicists make a distinction between trust and
trustworthiness (Baier 1986). Trust is a verb: placing one's trust in
someone else, with respect to some activity or domain. Trustworthiness
is an assessment, of whether or not that trust is appropriate. According
to the objection to the ideal of transparency, transparency would not
change how trustworthy science actually is --- the science is conducted
in the same way --- but might reduce public perceptions of
trustworthiness --- as members of the public see how the scientific
sausage is made. For instance, members of the public generally accept
the ideal of value-free science; values disclosures would violate this
ideal, making scientists (incorrectly) appear biased and untrustworthy
(John 2017; Kovaka 2021).

@@

After viewing the slide, the participant is then asked to rate
Dr.~Spence's trustworthiness using a semantic differential scale.
Elliott et al. (2017) used an ad hoc scale; we used the Muenster
Epistemic Trustworthiness Inventory (METI; Hendriks, Kienhues, and
Bromme 2015), which substantially overlaps but is not identical to the
Elliott et al. (2017) scale. Even though the ad hoc scale created by
Elliot and colleagues had acceptable internal reliability, the METI was
developed and psychometrically validated to measure the perceived
trustworthiness of experts. On the METI, participants rate the target
scientist (the fictional Dr.~Spence) on a 1-7 scale for 14 semantic
differential items, where each item is anchored at the ends by a pair of
words, such as competent-incompetent or responsible-irresponsible. The
METI scale captures perceived trustworthiness of a target along three
dimensions: competence, benevolence, and integrity. However, the three
dimensions were strongly correlated in our sample (Pearson r values
ranging from .75 to .92). To avoid issues of multicollinearity in our
analyses, we averaged participant scores into a single composite
measure. In both Elliott et al. (2017) and our replication analysis, all
items were averaged together to form a 1-7 composite measure of
trustworthiness. To aid interpretation, in the analysis the direction of
the scale has been set so that increased values correspond to increased
perceived trustworthiness.

As its name indicates, METI assesses perceived trustworthiness rather
than trust (such as accepting Dr.~Spence's conclusion about BPA). It is
conceivable that participants might trust Dr.~Spence's conclusion even
if they judge Spence to be untrustworthy, or vice versa. But we would
typically trust and trustworthiness to go together.

@@

\begin{description}
\tightlist
\item[H1. Modest correlation between values and ideology]
(a) Political liberals are more likely to prioritize public health over
economic growth, compared to political conservatives; but (b) a majority
of political conservatives prioritize public health.
\item[H2. Consumer risk sensitivity]
Scientists who find that a chemical harms human health are perceived as
more trustworthy than scientists who find that a chemical does not cause
harm.
\item[H3. Transparency penalty]
Scientists who disclose values are perceived as less trustworthy than
scientists who do not.
\item[H4. Shared values]
Given that the scientist discloses values, if the participant and the
scientist share the same values, the scientist is perceived as more
trustworthy than if the participant and scientist have discordant
values.
\item[H5. Variation in effects]
The magnitude of the effects for hypotheses 2-4 vary depending on
whether the participant prioritizes public health or economic growth.
\end{description}

@@

Hypothesis 1 was analyzed using Spearman rank order correlation to test
H1a and a visual inspection of descriptive statistics to test H1b.
Hypotheses 2-5 were analyzed using linear regression models as a common
framework, with a direct acyclic graph (DAG) constructed \emph{a priori}
to identify appropriate adjustments (covariates) for H4 and 5. Because
Elliott et al. (2017) made their data publicly available, and our
analytical approaches differ somewhat from theirs, we conducted parallel
analyses of their data, and report these parallel results for H2 and 3.
Exploratory data analysis was used throughout to support data validation
and aid interpretation.

@@

After excluding participants who declined consent after opening the
survey or did not complete the survey, we had 988 participants in the
full analysis sample (\(M_{age}\) = 44-years-old, \(SD_{age}\) =
16-years, Woman/Female = 498, Man/Male = 458, White = 712, Black = 124,
Asian or Pacific Islander = 63, Hispanic = 33, American Indian or
Alaskan Native = 5, Mixed or Other = 51). Participants were randomly
assigned to condition, with 163 assigned to the No Disclosure + Causes
Harm condition, 165 assigned to the No Disclosure + Does Not Cause Harm
condition, 165 assigned to the Economic Growth + Causes Harm condition,
165 assigned to the Economic Growth + Does Not Cause Harm condition, 168
assigned to the Public Health + Causes Harm condition, and 162 assigned
to the Public Health + Does Not Cause Harm condition
(Table~\ref{tbl-condition}). Due to researcher error a question about
participants' values was not included in the original survey. Of the
full 988 participants, 844 participants (85\%) responded to the followup
question about their own values (participant prioritizes economic growth
or public health). Consequently, subsamples for hypotheses 4 and 5 were
substantially smaller than the full analysis sample.

@@

Elliott et al. (2017) claim evidence for a shared values effect by
comparing values disclosure vs.~no disclosure (that is, estimating a
transparency effect) in quasi-independent subsamples across a 2
\(\times\) 2 \(\times\) 2 design (participant values \(\times\)
scientist conclusion \(\times\) scientist values; Elliott et al. (2017)
fig.~1). This approach is difficult to interpret; their reasoning seems
to be that this estimate is statistically significant in 4/8 cells, and
in 3/4 of these cells (1 where respondent values economic growth) the
participant and scientist share values. But these analyses are
underpowered (for example, sample sizes are well below 100 for some of
the cells in which participants value economic growth) and the overall
approach cannot distinguish a shared values effect from other potential
effects (scientist values, participant values). Applying our regression
analysis approach to the data published by Elliott et al. (2017), the
estimate for shared values was not statistically significant, \(\beta\)
= 0.29, \emph{t}(335) = 1.75, \emph{p} = 0.08; while the estimate for
scientist values was statistically significant, \(\beta\) = 0.41,
\emph{t}(335) = 2.48, \emph{p} = 0.014. It seems likely that claims of a
shared values effect could have been an artifact of a less appropriate
analytic approach given the nature of the data.

The disagreement over a transparency penalty is more difficult to
explain. Using our approach and the data from Elliott et al. (2017), the
estimate for a values disclosure was statistically significant
(Table~\ref{tbl-bc} model 1). One possible explanation is that, over the
course of the COVID-19 pandemic, members of the general public have
become used to ``scientists'' (broadly including both bench researchers
and public health officials) making claims about the importance of
protecting public health. A values disclosure that might have been
regarded as violating the value-free ideal, pre-pandemic, might now be
seen as routine.

@@

All together, regarding the philosophical debates over values in
science, the evidence from this replication study suggests that the
effects of transparency on perceived trustworthiness might be
context-dependent. Transparency about the influence of might be
detrimental in some cases, beneficial in others, and neutral in still
others. Notably, both of the papers cited as objecting to transparency
(John 2017; Kovaka 2021) focus on climate change, while many of
Elliott's examples come from toxicology or environmental public health.
The critics might be right that transparency has undermined trust in the
climate context, while Elliott might be right that transparency can be
neutral or even beneficial in some public health contexts. Because
philosophers of science tend to work with a small number of rich case
studies, methods such as qualitative comparative analysis (QCA; Ragin
2013) would be worth exploring.

An obvious difference between climate and BPA is the public prominence
and political polarization of the two issues. For decades, the fossil
fuels industry has used a range of strategies to successfully attack the
perceived trustworthiness of climate scientists, especially in the minds
of conservative publics (Oreskes and Conway 2011; Supran and Oreskes
2017). While the chemical industry has also had a substantial influence
on the environmental safety regulatory system (Michaels 2008; Vogel
2013), controversies over BPA have received much less public attention.
A LexiNexis search indicates that uses of ``bisphenol'' in US newspapers
rose and then almost immediately fell around 2008, shortly after a US
consensus conference concluded that ``BPA at concentrations found in the
human body is associated with organizational changes'' in a number of
major physiological systems (vom Saal et al. 2007).

In more prominent and/or polarized cases than BPA, a number of other
psychological mechanisms might be activated. Motivated reasoning or
identity-protective cognition (Kahan et al. 2012) might cause members of
the general public to rate a scientist as more or less trustworthy
insofar as the scientist agrees with the respondent's ``side'' of the
debate, regardless of whether and which values are disclosed. That is,
the scientist conclusion might have a larger effect than scientist
values. Alternatively, respondents might use values disclosures to
identify which scientists are on ``their side,'' and so scientist values
would have a larger effect than scientist conclusion. Careful attention
to experimental design will be needed to separate these potential
multiway interactions between conclusions, disclosures, and public
prominence.

@@

\leavevmode\vadjust pre{\hypertarget{ref-KahanPolarizingImpactScience2012}{}}%
Kahan, Dan M, Ellen Peters, Maggie Wittlin, Paul Slovic, Lisa Larrimore
Ouellette, Donald Braman, and Gregory Mandel. 2012. {``The Polarizing
Impact of Science Literacy and Numeracy on Perceived Climate Change
Risks.''} \emph{Nature Climate Change} 2 (10): 732--35.
\url{https://doi.org/10.1038/nclimate1547}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-MichaelsDoubtTheirProduct2008}{}}%
Michaels, David. 2008. \emph{Doubt is their product how industry's
assault on science threatens your health}. Oxford: Oxford Univ. Press.

@@

\leavevmode\vadjust pre{\hypertarget{ref-OreskesMerchantsDoubtHow2011}{}}%
Oreskes, Naomi, and Erik M. Conway. 2011. \emph{Merchants of Doubt: How
a Handful of Scientists Obscured the Truth on Issues from Tobacco Smoke
to Global Warming}. New York, NY, USA: Bloomsbury.

@@

\leavevmode\vadjust pre{\hypertarget{ref-RaginComparativeMethodMoving2013}{}}%
Ragin, Charles C. 2013. \emph{The comparative method: Moving beyond
qualitative and quantitative strategies}. Oakland, California:
University of California Press.

@@

\leavevmode\vadjust pre{\hypertarget{ref-vomSaalChapelHillBisphenol2007}{}}%
Saal, Frederick S. vom, Benson T. Akingbemi, Scott M. Belcher, Linda S.
Birnbaum, D. Andrew Crain, Marcus Eriksen, Francesca Farabollini, et al.
2007. {``Chapel Hill Bisphenol A Expert Panel Consensus Statement:
Integration of Mechanisms, Effects in Animals and Potential to Impact
Human Health at Current Levels of Exposure.''} \emph{Reproductive
Toxicology} 24 (2): 131--38.
\url{https://doi.org/10.1016/j.reprotox.2007.07.005}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-SupranAssessingExxonMobilClimate2017}{}}%
Supran, Geoffrey, and Naomi Oreskes. 2017. {``Assessing ExxonMobil's
Climate Change Communications (1977--2014).''} \emph{Environmental
Research Letters} 12 (8): 084019.
\url{https://doi.org/10.1088/1748-9326/aa815f}.

\leavevmode\vadjust pre{\hypertarget{ref-VogelItSafeBPA2013}{}}%
Vogel, Sarah A. 2013. \emph{Is It Safe?: BPA and the Struggle to Define
the Safety of Chemicals}. Oakland, California: University of California
Press.
